---
title: 'STA Homework 6 1.'
author: "Chen Zihao 915490404"
date: "2018年6月3日"
output: pdf_document
---
### 1.

(a)

$$
\rho_{Y,Z} = \frac{cov(X,Y)}{\sigma_X\sigma_Y}=\frac {E[(X-\mu_X)(Y-\mu_Y)]}{\sigma_X\sigma_Y}
$$
Use Pearson correlation coefficient as the estimate of the correlation coefficient.

$$
r=\frac {\sum_{i=1}^n(X_i-\bar X)(Y_i-\bar Y)}{\sqrt {\sum_{i=1}^n(X_i-\bar X)^2}\sqrt{\sum_{i=1}^n(Y_i-\bar Y)^2}}
$$


```{r}
Y = c(576,635,558,578,666,580,555,661,651,605,653,575,545,572,594)
Z = c(3.39,3.30,2.81,3.03,3.44,3.07,3.00,3.43,3.36,3.13,3.12,2.74,2.76,2.88,3.96)
```

(b)

Estimate the standard error of the Pearson correlation coefficient.

using jackknife
$$
\hat {se}_{jack}=\{\frac {n-1}n\sum[\hat r_{(i)}-\hat r_{(.)}]^2\}^{1/2}
$$

we can get $\hat {se}_{jack}$ is
```{r}
r = c()
n = length(Y)
for (i in 1:n){
  r[i]=cor(Y[-i],Z[-i])
}
se_jack = ((n-1)/n*sum((r-mean(r))^2))^(1/2)
se_jack
```

using bootstrap

1. select B independent bootstrap samples
2. calculate the bootstrap replication corresponding to each bootstrap samples.
3. estimate the S.E. by the sample standard deviation of the B replicates.

we can get $\hat {se}_{B}$ is
```{r}
set.seed(2018)
r = c()
B = 200
for (i in 1:B){
  sample = floor(runif(length(Y),1,length(Y)+1))
  r[i]=cor(Y[sample],Z[sample])
}

se_B = (sum((r-mean(r))^2)/(B-1))^(1/2)
se_B
```

(c)

1. using "normal theory"

$\hat \theta_{(.)}\pm \Phi(0.975)\hat {se}_B$

the C.I is 
```{r}
c(mean(r)-qnorm(0.975)*se_B,mean(r)+qnorm(0.975)*se_B)
```

2. using bootstrap $t$-interval approaches.

1. generate B bootstrap samples 
2. for each sample, calculate the Pearson correlation coefficient and estimated standard error (using bootstrap in the bootstrap, as in (b) to get the estimated standard error).
3. calculate $z^*(b)=\frac {\hat \theta(b)-\hat\theta}{\hat{se}(b)}$ for each b.
4. get the $\alpha$ percentile of $z^*(b)$ is estimated by the value $\hat t ^{(\alpha)}$ such that $\#\{z^*(b)<\hat t ^{(\alpha)}\}/B=\alpha$
5. the bootstrap-t confidence interval is $(\hat\theta-\hat t^{(1-\alpha)}\hat{se},(\hat\theta-\hat t^{(\alpha)}\hat{se})$,where $\hat{se}$ is the standard deviation of $\hat\theta^*(b)$'s.
```{r}
B=1000
theta = c()
se = c()
for(i in 1:B){
  sample = ceiling(runif(length(Y),0,length(Y)))
  sample_Y = Y[sample]
  sample_Z = Z[sample]
  theta[i] = cor(sample_Y,sample_Z)
  r = c()
  for (j in 1:200){    
    sample2 = ceiling(runif(length(Y),0,length(Y)))
    r[j] = cor(sample_Y[sample2],sample_Z[sample2])
  }
  se[i] = sd(r)
}

z=c()
for (i in 1:length(se)){
  z[i]=(theta[i]-mean(theta))/se[i]
}
c(mean(theta)-sort(z)[0.975*B]*sd(theta),mean(theta)-sort(z)[0.025*B]*sd(theta))
```
