---
title: 'STA Homework 6 3.1'
author: "Chen Zihao 915490404"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,fig.height=3)
```

For the genetic algorithm, I just copy the code from my homework assignment 2 here.

Here is the result for the test function in Assignment 2.

The true plot.
```{r}
truefunction<-function(x){
  t <- c(0.1, 0.13, 0.15, 0.23, 0.25, 0.4, 0.44, 0.65, 0.76, 0.78, 0.81)
  h <- c(4, -5, 3, -4, 5, -4.2, 2.1, 4.3, -3.1, 2.1, -4.2)
  temp <- 0
  for(i in 1:11) {
    temp <- temp + h[i]/2 * (1 + sign(x - t[i]))
    }
  return(temp)
  }
n<-512
x<-(0:(n-1))/n
f<-truefunction(x)
set.seed(0401)
y<-f+rnorm(f)/3
plot(x,y)
lines(x,f)
```

```{r}
S=300
P=0.9
Pc=0.05
Nsame=20

n = length(x)
chromosomes1 = matrix(rbinom(n*S,1,0.5),S,n)
chromosomes1[,1] = 0
```

```{r,echo=FALSE}
pred = function (chrom,y){
  bj = which (chrom == 1) #b1 to b_{B-1}
  B = sum(chrom)+1
  nhat = bj[-1]-bj[-B+1]
  nhat = c(bj[1]-1,nhat,length(y)-bj[B-1]+1)
  
  yhat = rep(mean(y[1:(bj[1]-1)]),nhat[1])
  
  for (i in 1:(B-2)) {
    yhat = c(yhat,rep(mean(y[bj[i]:(bj[i+1]-1)]),nhat[i+1]))
  }
  yhat = c(yhat,rep(mean(y[bj[(B-1)]:length(y)]),nhat[length(nhat)]))
  yhat
}

MDL = function(chrom,y){
  n = length(y)
  
  B = sum(chrom)+1 #the number of pieces
  
  #b_j where is the break points.
  bj=which(chrom == 1)
  
  #how many obervations in each pieces. 
  #Attention for the case when the first element of the chromosome is 1. 
  #if it is the case, then the first piece will contain 0 observations in it.
  nhat = bj[-1]-bj[-B+1]
  nhat = c(bj[1]-1,nhat,length(y)-bj[B-1]+1)
  if (is.na(nhat[1])){
    MD=NA
  } else {
    MD = B*log(n)+sum(log(nhat))/2+n/2*log(sum((y-pred(chrom,y))^2)/n)
    }
  MD
}

AIC = function(chrom,y){
  B = sum(chrom)+1
  n = length(y)
  n*log(sum((y-pred(chrom,y))^2)/n)+log(n)*2*B
}

value = function (chrom,y,method){
  v = rep(0,nrow(chrom))
  for (i in 1:nrow(chrom)){
    if (method == 1) {
      v[i] = MDL(chrom[i,],y)
    } else if (method == 2){
      v[i] = AIC(chrom[i,],y)
    }
  }
  v
}
```

```{r,echo=FALSE}
Sort = function (chrom,y,method){
  chrom[order(value(chrom,y,method),decreasing = TRUE),]
}
```

```{r,echo=FALSE}
crossover = function(chrom){
  n = nrow(chrom)
  p = ncol(chrom)
  #the input should be ordered
  
  #select the parents
  i1=ceiling(sqrt(2*ceiling(runif(1,0,n*(n+1)/2))+0.25)-0.5)
  i2=ceiling(sqrt(2*ceiling(runif(1,0,n*(n+1)/2))+0.25)-0.5)
  
  a=chrom[i1,]
  b=chrom[i2,]
  
  k = ceiling(runif(1,1,p))# 1<k<=p
  
  c(a[1:k-1],b[k:p])
}

mutation = function(chrom,Pc){
  n = nrow(chrom)
  i1=ceiling(sqrt(2*ceiling(runif(1,0,n*(n+1)/2))+0.25)-0.5)
  a=chrom[i1,]
  for (i in 2:ncol(chrom)){
    k = rbinom(1,1,Pc)
    if (k == 1){
      a[i]=1-a[i]
    }
  }
  a
}
```

```{r,echo=FALSE}
offspring = function(chrom,p,pc){
  offs = chrom[nrow(chrom),] # keep the best chrom in the previous gen.
  for (i in 2:nrow(chrom)){
    j = rbinom(1,1,p)
    if (j==1){
      offs = rbind(offs,crossover(chrom))
    } else{
      offs = rbind(offs,mutation(chrom,pc))
    }
  }
  offs
}
```

```{r,echo=FALSE}
GA = function (y,method){
  S=300
  p=0.9
  pc=0.05
  Nsame=20
  
  n = length(y)
  chrom = matrix(rbinom(n*S,1,0.5),S,n)
  chrom[,1] = 0
  chrom = Sort(chrom,y,method)
  
  best = chrom[S,]
  t=0
  
  while (t<Nsame){
    previous = best
    chrom = offspring(chrom,p,pc)
    chrom = Sort(chrom,y,method)
    best = chrom[S,]
    if (all(previous == best)){
      t = t+1
    }else {t=0}
  }
  best
}
```

```{r,echo=FALSE}
#to make it a R function to make the out puts
final = function(y,method){
  #method: 1 is using MDL and 2 is using AIC
  #y : the noise data
  results = GA(y,method)
  if (method == 1){
    method = "MDL"
  } else {
    method = "AIC"
  }
  plot(x,y,main = paste("Genetic Algorithms with",method))
  lines(x,pred(results,y))
}
```

I just use the MDL method to generate the y and the plot.

```{r,echo=FALSE}
results = GA(y,1)
plot(x,y,main = "Genetic Algorithms with MDL")
yhat = pred(results,y)
lines(x,yhat)
```

for the sencond model

```{r}
f2 = (4*x-2)+2*exp(-16*(4*x-2)^2)
noise = c()
for (i in 1:n){
  noise[i] = rnorm(1,0,abs(f2[i]/5))
}
y2 = f2+noise
plot(x,y2,cex = 0.3,main = "True line and the dot plot")
lines(x,f2)
```

```{r,echo=FALSE}
results2 = GA(y2,1)
plot(x,y2,main = "Genetic Algorithms with MDL",cex = 0.3)
yhat2 = pred(results2,y2)
lines(x,yhat2)
```

```{r}
func = function(y){pred(GA(y,1),y)}
```

(b)

part 1

Get the residuals of the model and resample the residuals then obtain the bootstrap sample.

```{r}
boot_re = function(yhat,y,B){
  #get the residuals
  residual = y - yhat
  #resample the residuals ans get the bootstrap sample
  r = matrix(0,B,length(y))
  for (i in 1:B){
    sample = ceiling(runif(length(y),0,length(y)))
    r[i,] = y + residual[sample]
  }
  return(r)
}

boot_sample1 = boot_re(yhat,y,1000)
boot_sample2 = boot_re(yhat2,y2,1000)
```

and then run the algorithm to get 200 times bootstrap predicted value.

```{r}
library(snow)
cl = makeCluster(7)
clusterExport(cl,c("pred","GA","Sort","value","MDL","offspring","crossover","mutation") )
result = parApply(cl,boot_sample1,1,func)
result2 = parApply(cl,boot_sample2,1,func)
stopCluster(cl)
```

consturct pointwise C.I. and plot it like below

```{r}
k = 100
CI1 = result[,c(1,2)]
CI2 = result2[,c(1,2)]
for (i in 1:n){
  CI1[i,] = c(yhat[i]-qnorm(0.975)*sd(result[i,]),yhat[i]+qnorm(0.975)*sd(result[i,]))
  CI2[i,] = c(yhat2[i]-qnorm(0.975)*sd(result[i,]),yhat2[i]+qnorm(0.975)*sd(result[i,]))
  #CI1[i,]=sort(result[i,])[c(max(1,floor(0.025*k)),max(1,ceiling(0.975*k)))]
  #CI2[i,]=sort(result2[i,])[c(max(1,floor(0.025*k)),max(1,ceiling(0.975*k)))]
}
```

```{r,fig.height=6}
plot(x,y,main = "Genetic Algorithms with MDL",cex = 0.3)
lines(x,CI1[,1])
lines(x,CI1[,2])
lines(x,f,col="red")
```

```{r,fig.height=6}
plot(x,y2,main = "Genetic Algorithms with MDL",cex = 0.3)
lines(x,CI2[,1])
lines(x,CI2[,2])
lines(x,f2,col="red")
```

part 2 

bootstrapping pairs

resample pairs from the original data and then get the predicted values and consturcted the CI pointwise. plot it like below

```{r}
boot_pairs = function(x,y,B){
  
  order_boot = matrix(0,B,length(y))
  y_boot = matrix(0,B,length(y))
  for (i in 1:B){
    #get the random index
    sample = sort(ceiling(runif(length(y),0,length(y))))
    #get the sample
    order_boot[i,] = sample
    y_boot[i,] = y[sample]
  }
  return(list(order=order_boot,y=y_boot))
}
bp = boot_pairs(x,y,1000)
bp1o = bp$order
bp1y = bp$y
bp2 = boot_pairs(x,y2,1000)
bp2o = bp2$order
bp2y = bp2$y
```


```{r}
cl = makeCluster(7)
clusterExport(cl,c("pred","GA","Sort","value","MDL","offspring","crossover","mutation") )
result3 = parApply(cl,bp1y,1,func)
result4 = parApply(cl,bp2y,1,func)
stopCluster(cl)
```

```{r}
CI3 = matrix(0,n,2)
CI4 = matrix(0,n,2)
for (i in 1:n){
  a = result3[bp1o == i]
  CI3[i,] = c(yhat[i]-qnorm(0.975)*sd(a),yhat[i]+qnorm(0.975)*sd(a))
  a = result4[bp1o == i]
  CI4[i,] = c(yhat2[i]-qnorm(0.975)*sd(a),yhat2[i]+qnorm(0.975)*sd(a))
  # a = sort(result3[bp1o == i])
  # k = length(a)
  # CI3[i,]=a[c(max(1,floor(0.025*k)),max(1,ceiling(0.975*k)))]
  # 
  # a = sort(result4[bp2o == i])
  # k = length(a)
  # CI4[i,]=a[c(max(1,floor(0.025*k)),max(1,ceiling(0.975*k)))]
}
```

```{r,fig.height=6}
plot(x,y,main = "Genetic Algorithms with MDL",cex = 0.3,ylim = c(min(CI3),max(CI3)))
lines(x,CI3[,1])
lines(x,CI3[,2])
lines(x,f,col="red")
```

```{r,fig.height=6}
plot(x,y2,main = "Genetic Algorithms with MDL",cex = 0.3,ylim = c(min(CI4),max(CI4)))
lines(x,CI4[,1])
lines(x,CI4[,2])
lines(x,f2,col="red")
```

```{r}
boot_pairs2 = function(x,y,B){
  
  order_boot = matrix(0,B,length(y))
  y_boot = matrix(0,B,length(y))
  for (i in 1:B){
    #get the random index
    ord = unique(ceiling(runif(length(y),0,length(y))))
    sample = sort(ord)
    #get the sample
    
    for (j in 1:length(sample)){
      order_boot[i,j] = sample[j]
      y_boot[i,j] = y[sample[j]]
    }
  }
  return(list(order=order_boot,y=y_boot))
}
bp_2 = boot_pairs2(x,y,98*4)
bp1o2 = bp_2$order
bp1y2 = bp_2$y
bp2_2 = boot_pairs2(x,y2,98*4)
bp2o2 = bp_2$order
bp2y2 = bp_2$y
```


```{r}
cl = makeCluster(7)
clusterExport(cl,c("pred","GA","Sort","value","MDL","offspring","crossover","mutation") )
result5 = parApply(cl,bp1y2,1,func)
result6 = parApply(cl,bp2y2,1,func)
stopCluster(cl)
```

```{r}
CI5 = matrix(0,n,2)
CI6 = matrix(0,n,2)
for (i in 1:n){
  # a = result3[bp1o == i]
  # CI3[i,] = c(yhat[i]-qnorm(0.975)*sd(a),yhat[i]+qnorm(0.975)*sd(a))
  # a = result4[bp1o == i]
  # CI4[i,] = c(yhat2[i]-qnorm(0.975)*sd(a),yhat2[i]+qnorm(0.975)*sd(a))
  a = sort(result5[bp1o2 == i])
  k = length(a)
  CI5[i,]=a[c(max(1,floor(0.025*k)),max(1,ceiling(0.975*k)))]

  a = sort(result6[bp2o2 == i])
  k = length(a)
  CI6[i,]=a[c(max(1,floor(0.025*k)),max(1,ceiling(0.975*k)))]
}
```

```{r,fig.height=6}
plot(x,y,main = "Genetic Algorithms with MDL",cex = 0.3)#,ylim = c(min(CI5),max(CI5)))
lines(x,CI5[,1])
lines(x,CI5[,2])
lines(x,f,col="red")
```

```{r,fig.height=6}
plot(x,y2,main = "Genetic Algorithms with MDL",cex = 0.3,ylim = c(min(CI6,na.rm = TRUE),max(CI6,na.rm = TRUE)))
lines(x,CI6[,1])
lines(x,CI6[,2])
lines(x,f2,col="red")
```