---
title: 'STA Homework 6 2.'
author: "Chen Zihao 915490404"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,fig.height=3)
```

### 2.

(a) Find the distribution.

$$
\begin{split}
P(X_{\max}<x) &= \prod_{i=1}^{n}P(X_i<x)\\
& = (x/\theta)^n\\
f_{X_{\max}}(x)&=\frac {nx^{n-1}} {\theta^n}
\end{split}
$$

so that we get the pdf of $X_{\max}$

(b) Derive the analytic expression for the variance.

$$
\begin{split}
E(\hat\theta)&=\int_0^\theta x\frac {nx^{n-1}} {\theta^n}dx\\
& = \frac n {\theta^n}\times\frac 1 {n+1} x^{n+1}|_0^\theta\\
& = \frac n {n+1}\theta\\
E(\hat\theta^2)&=\int_0^\theta x^2\frac {nx^{n-1}} {\theta^n}dx\\
& = \frac n {\theta^n}\times \frac 1 {n+2}x^{n+2}|_0^\theta\\
& = \frac n {n+2}\theta^2\\
Var(\hat\theta) &= E(\hat\theta^2)-E(\hat\theta)^2\\
 & = \frac n {n+2}\theta^2-(\frac n {n+1}\theta)^2\\
 & = \frac {n\theta^2}{(n+1)^2(n+2)}
\end{split}
$$

(c) Generate a data set of size $n = 50$ and $\theta = 3$. Then generate $B = 5000$ bootstrap samples using parametric bootstrap. Use the bootstrap samples to approximate $Var_{F_\theta}(\hat\theta)$. Compare your answer to (b).

Take the maximum of the sample as the $\hat\theta$, simulate bootstrap samples from unif(0,$\hat \theta$)

the parametric bootstrap result is 

```{r}
set.seed(201806)
X = runif(50,0,3)
r1=c()
for (i in 1:5000){
  sample = runif(length(X),0,max(X))
  r1[i] = max(sample)
}
var(r1)
```


the answer to (b) is 
```{r}
50*3^2/((50+1)^2*(50+2))
```

They are close.

(d)

the nonparametric bootstrap samples results is
```{r}
r2=c()
for (j in 1:5000){    
  sample = ceiling(runif(length(X),0,length(X)))
  r2[j] = max(X[sample])
}
var(r2)
```

(e)

```{r}
hist(r1,main = "The parametric bootstrap",xlab = "theta hat *")
hist(r2,main = "The non-parametric bootstrap",xlab = "theta hat *")
```

(f)

the true distribution of $\hat\theta$


```{r}
r3 = c()
for (j in 1:5000){    
  sample = runif(length(X),0,3)
  r3[j] = max(sample)
}
hist(r3,main = "The 5000 samples of the theta hat from unif(0,3)",xlab = "theta hat")
```

The non-parametric plot did not look like the true one.

The reason is that for each non-parametric bootstrap, the chance of not choosing the largest value of the sample as the $\hat\theta$ is (49/50)^50 = 0.3641697 which is very low.

