---
title: "STA 243 Assignment 2"
author: "Chen Zihao 915490404"
output:
  pdf_document: default
---
# 1. Simulated Annealing

(Chen zihao:75%, Jichen Xu:25%)

 Let input the distance matrix into R first by pasted from the pdf. 
```{r,echo=FALSE}
#Distance matrix
D = matrix (c(
0,1,2,4,9,8,3,2,1,5,7,1,2,9,3,
1,0,5,3,7,2,5,1,3,4,6,6,6,1,9,
2,5,0,6,1,4,7,7,1,6,5,9,1,3,4,
4,3,6,0,5,2,1,6,5,4,2,1,2,1,3,
9,7,1,5,0,9,1,1,2,1,3,6,8,2,5,
8,2,4,2,9,0,3,5,4,7,8,3,1,2,5,
3,5,7,1,1,3,0,2,6,1,7,9,5,1,4,
2,1,7,6,1,5,2,0,9,4,2,1,1,7,8,
1,3,1,5,2,4,6,9,0,3,3,5,1,6,4,
5,4,6,4,1,7,1,4,3,0,9,1,8,5,2,
7,6,5,2,3,8,7,2,3,9,0,2,1,8,1,
1,6,9,1,6,3,9,1,5,1,2,0,5,4,3,
2,6,1,2,8,1,5,1,1,8,1,5,0,9,6,
9,1,3,1,2,2,1,7,6,5,8,4,9,0,7,
3,9,4,3,5,5,4,8,4,2,1,3,6,7,0
),15,15)
```

1.1 First, we need to define the path in a vector form. label the city with numbers instead of alphabet, e.g. 1 for A , 2 for B, etc. then we can get the path which is also the candidate in this problem:

$\theta = (i_1,...,i_{15})$ where $i_j\ne i_k$ and $i_j\in \{1,2,...,15\}$

the candidate space contains all the possible $\theta$.

To solve the travel problem, we should also compute the total distance. Here is the function to calculate the total distances for each candidate path.

$$
Distance(\theta) = d(i_{15},i_{1})+\sum_{j=1}^{14}d(i_j,i_{j+1})
$$

where $d(i_{j},i_{k})$ is the distance from $i_{j}$ city to the $i_{k}$ city which is given in the matrix mentioned in the beginning  $d(i_{j},i_{k})=D_{jk}$ 

```{r,echo=FALSE}
Distance = function(x){
  dis = D[x[15],x[1]]
  for (i in 1:14){
    dis = dis + D[x[i],x[i+1]]
  }
  dis
}
```

1.2 Second, the Simulated Anealing algorithm part.

1.2.0 Initialization

initialize $\tau_1= 400$ and other things mentioned in the question.

1.2.1 step 1: sample a candidate

From the first part, we have define what is the candidate space. To use a uniform distribution for the proposal density. Get the random number j and k uniformly without replacement and then  switch $i_j$ and $i_k$ and I did it twice to get the neighbour which means in this step I will exchange at least 2 at most 4 variables.

```{r,echo=FALSE}
step1 = function (best) {
  order = sample(1:15,2)
  
  theta = best
  temp = theta[order[1]]
  theta[order[1]] = theta[order[2]]
  theta[order[2]] = temp
  theta
}
```

1.2.2 step 2: calculate the distance and compare them.

a) if the new candidate is better than the previous best one, take the candidate as the best.

$\Delta=Distance(\theta^*)-Distance(\theta_k)<0$ where $\theta_k$ is the recent best solution and $\theta^*$ is the new candidate drawn by the step 1, then take $\theta^*$ as $\theta_{k+1}$

b) Otherwise, $\theta_{k+1}=\theta^*$ with a probability $\exp(-\frac\Delta {\tau_j})$, keep $\theta_{k+1}=\theta_{k}$ o.w.

```{r,echo=FALSE}
step2 = function (theta,best,tau){
  Delta = Distance(theta) - Distance(best)
  if (Delta > 0){
    p = exp(-Delta/tau)
    coin = rbinom(1,1,p)
    if(coin == 1){
      best = theta
    }
  }
  else best = theta
  best
}
```

1.2.3 step 3: repeat step 1 and 2 $m_j=100$ times.

1.2.4 step 4: update $\tau_{j+1}=\alpha(\tau_j), m_{j+1}=\beta(m_j)$ and move to stage j+1.

1.2.5 step 5: reheat.

run the algorithm again with the initial point as the previous result.

here is the result for different parameters:

1. $p=0.999,tau = 400$

the result path is of length:
```{r,echo = FALSE}
#the algorithm
sa = function (best){
  for (j in 1:1000){#run 1000 stages
    for (i in 1:m){
      theta = step1(best)#step 1 part 1 switch once
      theta = step1(theta)# step 1 part 2 switch second time
      best = step2(theta,best,tau) # step 2
      }#run for m times , step 3
    tau = p*tau #step 4
    }
  best
}
```


```{r,echo = FALSE}
#initialization
best = 1:15#the initial point,sample(1:15) will be the random initial point.
p=0.999
m=100
tau = 400

#run and reheat it,
# for (k in 1:3){
   best = sa(best)
# }
# I did not reheat it in this part because it is useless. I save some output time when i knit it as the pdf.
Distance(best)
```
the result is bad because $exp(-30/(400*0.999^{1000}))=0.815$, the chance to accept the worse case is still high even in the final stage. The temperture is still high in the end.

2. $p=0.99,tau = 400$
```{r,echo = FALSE}
#initialization
best = 1:15#the initial point,sample(1:15) will be the random initial point.
p=0.99
m=100
tau = 400

#run and reheat it
for (k in 1:3){
  best = sa(best)
}
Distance(best)
```

I run it several times, the results are stable around 17 to 20, which are good enough. One of the best path (with distance 17) I found is (3,9,1,2,8,12,10,15,11,13,6,14,4,7,5)

