---
title: "STA 243 Assignment 1"
author: "Chen Zihao 915490404"
output:
  pdf_document: default
  html_document: default
---
# 1. The Cauchy.....
##(a) 

the likelihood function is that

$$
L(\theta|x)=f_n(x|\theta)=\Pi_{i=1}^n\frac 1 {\pi[1+(x-\theta)^2]}
$$

then we can get:

1) the log likelihood function is that

$$
l(\theta)=-nlog\pi-\sum_{i=1}^n\log[1+(\theta-x_i)^2]
$$

2) take the first derivative and we can get

$$
l'(\theta)=-\sum_{i=1}^n\frac {2(\theta-x_i)}{1+(\theta-x_i)^2}
$$

3) take the second derivative and we can get

$$
l''(\theta)=-2\sum_{i=1}^n\frac {1+(\theta-x_i)^2-2(\theta-x_i)(\theta-x_i)}{[1+(\theta-x_i)^2]^2}=-2\sum_{i=1}^n\frac {1-(\theta-x_i)^2}{[1+(\theta-x_i)^2]^2}
$$

## (b) Show that the Fisher information is $I(\theta)=\frac n 2$

$$
\begin{split}
I(\theta)=-E(l''(\theta))&=2nE(\frac {1-(\theta-x_1)^2}{[1+(\theta-x_1)^2]^2})\\
&=\frac{2n}\pi\int_{-\infty}^\infty\frac 1 {[1+(x-\theta)^2]^2}-\frac {2(x-\theta)^2}{[1+(x-\theta)^2]^3} dx\\
&=\frac{2n}\pi\int_{-\infty}^\infty\frac 1 {[1+x^2]^2}-\frac {2x^2}{[1+x^2]^3} dx\\
&=\frac{2n}\pi\int_{-\infty}^\infty\frac 1 {[1+x^2]^2}-\frac {2}{[1+x^2]^2}+\frac {2}{[1+x^2]^3} dx\\
&=\frac{2n}\pi\int_{-\infty}^\infty\frac {-1} {[1+x^2]^2}+\frac {2}{[1+x^2]^3} dx\\
\end{split}
$$
let 
$$
F_k=\int_{-\infty}^\infty\frac 1 {[1+x^2]^k}dx
$$

we can get

$$
\begin{split}
F_k&=\int_{-\infty}^\infty\frac 1 {[1+x^2]^k}dx\\
&=\int_{-\infty}^\infty\frac {1+x^2} {[1+x^2]^{k+1}}dx\\
&=F_{k+1}+\int_{-\infty}^\infty\frac {2kx} {[1+x^2]^{k+1}}\frac x {2k}dx\\
&=F_{k+1}+\frac 1 {2k}\int_{-\infty}^\infty\frac {1} {[1+x^2]^{k}}dx=F_{k+1}+\frac 1 {2k} F_k\\
\end{split}
$$
By an integration by parts.

Hence we get

$F_1=\pi$ and $F_{k+1}=\frac{2k-1}{2k}F_k,k>1$

so that $F_1=\pi,F_2=\pi/2,F_3=3\pi/8$
$$
I(\theta)=\frac {2n}\pi [-F_2+2F_3]=\frac n 2
$$

## (c) Use the following data, graph the log likelihood function....

```{r,echo=FALSE}
x1=c(-13.87,-2.53,-2.44,-2.40,-1.75,-1.34,-1.05,-0.23,-0.07, 0.27, 1.77, 2.76, 3.29, 3.47, 3.71,3.80, 4.24, 4.53, 43.21, 56.75)

llf = function(theta){
  sapply(theta,function(theta){-length(x1)*log(pi)-sum(log(1+(theta-x1)^2))})
  }

t <- seq(from = -20 , to = 50 ,by = 0.01) 

plot(t,llf(t),type = "l",xlab = "theta", ylab="log likelihood function",main = "log likelihood function graph of the given data" )
```

It shows that the maximum is near 0,there is a local maximum and local minimum at little bit larger than 0, there seems a station point near 40.

## (d) Find the MLE for $\theta$ using the Newton-Raphson method....

```{r,echo=FALSE}
start=c(-11,-1, 0, 1.4, 4.1, 4.8, 7, 8, 38)
#first order
l1=function(x){
  -2*sum((x-x1)/(1+(x-x1)^2))
}
#second order
l2=function(x){
  -2*sum((1-(x-x1)^2)/(1+(x-x1)^2)^2)
}

#Newton-Raphson method
NR=function(x0){
  a=x0
  a1=a+2
  t=0
  while (abs(a1-a)>0.001){
    a1=a
    h=-l1(a)/l2(a)
    a=a+h
    if (t>30){return(a)}#maximum interation control
    if (abs(a1-a)>10^5){return(NA)}#it did not converge.
    t=t+1
  }
  a
}
```

with the given initial point, some points did not converge. The following table shows the results:

```{r,echo=FALSE}
a = sapply(start,NR)
rbind(start,"theta value" = a,"function value"=llf(a))
```

## (e) First use Fisher scoring to find the MLE for ....

First use Fisher scoring to find the MLE for $\theta$

```{r,echo = FALSE}
NR2=function(x0){
  a=x0
  a1=a+2
  t=0
  while (abs(a1-a)>0.001){
    a1=a
    h=2*l1(a)/length(x1)
    a=a+h
    if (t>30){return(a)}#maximum interation control
    if (abs(a1-a)>10^5){return(NA)}#if it did not converge.
    t=t+1
  }
  a
}
start2 = sapply(start,NR2)
```

the new initial point becomes
```{r,echo=FALSE}
start2
```

Then refine my estimate using Newton-Raphson.

```{r,echo=FALSE}
a = sapply(start2,NR)
rbind(start,"new initialpoint"=start2,"theta value" = a,"function value"=llf(a))
```

Compare with the previous one, Most of them is now stable although some of them are still trapped in the local maximum/minimum point. Those can not converage in (d) can converage now.