---
output: pdf_document
---

# 3.In chemical kinetics the Michaelis-Menten model...

```{r,echo=FALSE}
x = c(0.02, 0.06, 0.11, 0.22, 0.56, 1.10, 0.02, 0.06, 0.11, 0.22, 0.56, 1.10)
y = c(47, 97, 123, 152, 191, 200, 76, 107, 139, 159, 201, 207)
```

##(a) A quick way for finding rough estimates...

Using lm function in R, we can easily get the $\beta_0$ and $\beta_1$ with least squares 
```{r,echo=FALSE}
y2=1/y
x2=1/x
coef = lm(y2~x2)$coef
print(paste("beta0 is",coef[1],"beta1 is",coef[2]))
```

so that we can get $\hat\theta_1=1/\beta_0$ and $\hat\theta_2=\beta_1/\beta_0$ as follows:

```{r,echo=FALSE}
theta1=1/coef[1]
theta2=coef[2]/coef[1]
print(paste("theta1_hat is",theta1,"theta2_hat is",theta2))
```

## (b)

$$
\begin{split}
g(\theta_1,\theta_2)&=\sum_{i=1}^n(y_i-\frac{\theta_1x_i}{x_i+\theta_2})^2\\
\frac {\partial g} {\partial \theta_1}&=2\sum_{i=1}^n(y_i-\frac{\theta_1x_i}{x_i+\theta_2})(-\frac{x_i}{x_i+\theta_2})\\
\frac {\partial g} {\partial \theta_2}&= 2\sum_{i=1}^n(y_i-\frac{\theta_1x_i}{x_i+\theta_2})(\frac{\theta_1x_i}{(x_i+\theta_2)^2})\\
\frac {\partial^2 g} {(\partial \theta_1)^2}&= 2\sum_{i=1}^n(\frac{x_i}{x_i+\theta_2})^2\\
\frac {\partial^2 g} {\partial \theta_1\partial \theta_2}&= 2\sum_{i=1}^n(\frac{y_ix_i}{(x_i+\theta_2)^2}-\frac{2\theta_1x_i^2}{(x_i+\theta_2)^3})\\
\frac {\partial^2 g} {(\partial \theta_2)^2}&= 2\sum_{i=1}^n(\frac{-2y_i\theta_1x_i}{(x_i+\theta_2)^3}+\frac{3\theta_1^2x_i^2}{(x_i+\theta_2)^4})\\
\end{split}
$$

and then we can get the first derivative vector and the hessian matrix and do a Newton-Raphson algorithm. The code is in the code file, here is the results.

```{r,echo=FALSE}
g=function(theta){
  sum((y-(theta[1]*x)/(x+theta[2]))^2)
}
g1=function(theta){
  2*sum((y-(theta[1]*x)/(x+theta[2]))*(-x/(x+theta[2])))
}
g2=function(theta){
  2*sum((y-(theta[1]*x)/(x+theta[2]))*(theta[1]*x/(x+theta[2])^2))
}
g11=function(theta){
  2*sum((x/(x+theta[2]))^2)
}
g12=function(theta){
  2*sum(y*x/(x+theta[2])^2-2*theta[1]*x^2/(x+theta[2])^3)
}
g22=function(theta){
  2*sum(-2*y*x*theta[1]/(x+theta[2])^3+3*x^2*theta[1]^2/(x+theta[2])^4)
}

NR=function(theta){
  a=theta
  a1=a+2
  t=0
  while (sum(abs(a1-a))>0.0001){
    a1=a
    
    f1=c(g1(a),g2(a))
    hessian=matrix(c(g11(a),g12(a),g12(a),g22(a)),2,2)
    
    h=-solve(hessian)%*%f1
    a=a+h
    if (t>100){return(a)}#maximum interation control
    if (sum(abs(a1-a))>10^3){return(NA)}#it did not converge.
    t=t+1
  }
  a
}

theta=matrix(c(theta1,theta2),nrow=2)



```


```{r,echo=FALSE}
a=NR(theta)
cbind("theta1"=a[1,1],"theta2"=a[2,1])
```

## (c) Repeat (b) with the steepest descent algorithm.

change the update rule to
$$
X_{t+1}=X_t-\alpha_tg'(X_t)
$$
if $g(X_{t+1})>g(X_{t})$ then half $\alpha$

I firstly use $\alpha=1$, but it seems the magnitude of $\theta_1$ and $\theta_2$ is so different so that the algorithm will not converage to the ideal point. I try to set the $\alpha=diag(\theta_0)$ which will solve the magnitude problem, and the result is shown as below.

```{r,echo=FALSE}
sd=function(theta){
  a=theta
  a1=a+2
  t=0
  while (norm((a1-a),"2")>0.0001){
    t=t+1
    a1=a
    f1=c(g1(a),g2(a))
    
    alpha=theta
    a=a1-alpha*f1
    while (g(a)>g(a1)) {
      alpha=alpha/2
      a=a1-alpha*f1
    }
    
    if (t>10000){return(a)}#maximum interation control
    if (sum(abs(a1-a))>10^5){return(NA)}#it did not converge.
    
  }
  a
}
a=sd(theta)

cbind("theta1"=a[1,1],"theta2"=a[2,1])
```



## (d) Repeat (b) with the Gauss-Newton algorithm.

Consider 

$$
g(\theta)=-\sum_{i=1}^n(y_i-f_i(\theta))^2
$$
where
$$
\begin{split}
f_i(\theta)&=\frac{\theta_1 x_i}{x_i+\theta_2}\\
f'_i(\theta)&=[\frac{x_i}{x_i+\theta_2},-\frac{\theta_1x_i}{(x_i+\theta_2)^2}]^T\\
\end{split}
$$
then we get
$$
\begin{split}
A=A(\theta)&=[f_1'(\theta)^T,...,f_n'(\theta)^T]^T\\
Z=Z(\theta)&=[y_1-f_1(\theta),...,y_n-f_n(\theta)]^T\\
\end{split}
$$

The updating formula is 

$$
\theta_{t+1}=\theta_t+(A_t^TA_t)^{-1}A^T_tZ_t
$$

here is the results
```{r,echo=FALSE}
Z=function(theta){
  y-theta[1]*x/(x+theta[2])
}

A=function(theta){
  sapply(x,function(x){c(x/(x+theta[2]),-theta[1]*x/(x+theta[2])^2)})
}

gn=function(theta){
  a=theta
  a1=a+2
  t=0
  while (sum(abs(a1-a))>0.0001){
    a1=a
    a=a+solve(A(a)%*%t(A(a)))%*%A(a)%*%Z(a)
    if (t>100){return(a)}#maximum interation control
    if (sum(abs(a1-a))>10^3){return(NA)}#it did not converge.
    t=t+1
  }
  a
}
a=gn(theta)
cbind("theta1"=a[1,1],"theta2"=a[2,1])
```


