---
title: "STA 243 Assignment 4.3"
author: 
output:
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,fig.height=3)
```

## 3.  Control Variate Method

(a) folowing the instruction and here is the result.
```{r}
##3.(a)
U=runif(1500,0,1)
U_sum=0
for (i in 1:(length(U))) {
  U_sum=U_sum+1/(1+U[i])
}
I_MC=(1/1500)*(U_sum)
print(paste("the estimating integral value is",I_MC))
```
The exact value for I is $\ln 2 \approx 0.6931472$

```{r}
#variance
I_err=0
for (i in 1:(length(U))) {
  I_err=I_err+(1/(1+U[i])-I_MC)^2
}
I_var=1/(length(U)*(length(U)-1))*I_err

print(paste("the sample standard error of the estimating integral value is",sqrt(I_var)))
```
(b) and (c)

$E(1+U) = 1 + E(U) = 1+0.5 =1.5$.

$Var[c(U)] = Var(1+U) = \frac 1 {12}$ and $Cov[h(U),c(U)]=1-\frac {3\ln 2}{2}$

then we can get the best value for b, $b^*=\frac {Cov[h(U),c(U)]}{Var[c(U)]}\approx -0.4767$
```{r}
##(b)&(C)
C_sum=0
for (i in 1:(length(U))) {
  C_sum=C_sum+(1+U[i])
}
C=(1/1500)*(C_sum)
b = cov(1/(1+U), 1+U)/var(1+U)
I_CV=I_MC-b*(C-1.5)
print(paste("the estimating integral value is",I_CV))
#variance
I_err=0
for (i in 1:(length(U))) {
  I_err=I_err+(1/(1+U[i])-b*((1+U[i])-1.5)-I_CV)^2
}
I_var=1/(length(U)*(length(U)-1))*I_err
print(paste("the sample standard error of the estimating integral value is",sqrt(I_var)))

```

The variance of the CV version is less which implies the CV version is better.

(d)
Here we use $d(x)\sqrt x$ instead of $c(x) = 1+x$.

```{r}
##(d)
C_sum=0
for (i in 1:(length(U))) {
  C_sum=C_sum+(sqrt(U[i]))
}
C=(1/1500)*(C_sum)
b = cov(1/(1+U), sqrt(U))/var(sqrt(U))
I_CV=I_MC-b*(C-2/3)
print(paste("the estimating integral value is",I_CV))

#variance
I_err=0
for (i in 1:(length(U))) {
  I_err=I_err+(1/(1+U[i])-b*(sqrt(U[i])-2/3)-I_CV)^2
}
I_var=1/(length(U)*(length(U)-1))*I_err
print(paste("the sample standard error of the estimating integral value is",sqrt(I_var)))
```

the sample stanard error is even smaller than the previous one.