---
title: "STA 243 Assignment 4"
author: 
output:
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,fig.height=3)
```

## 6. Independence-Metropolis-Hastings Algorithm 

```{r,echo = FALSE}
#set gamma(k,theta)
#q1 and r1 using Gamma(k,x_{i-1}/k)
f = function(z){
  theta1 = 1.5
  theta2 = 2
  z^(-3/2)*exp(-theta1*z-theta2/z+2*sqrt(theta1*theta2)+log(sqrt(2*theta2)))
}

q1 = function(x,y,k){
  #y is the condition
  theta = y/k
  1/(gamma(k)*theta^k)*x^(k-1)*exp(-x/theta)
}

q2 = function(x,y,theta){
  #y is the condition
  k = y/theta
  1/(gamma(k)*theta^k)*x^(k-1)*exp(-x/theta)
}

r1 = function (x,y,k){
  min(1,f(y)*q1(x,y,k)/(f(x)*q1(y,x,k)))
}

r2 = function (x,y,theta){
  min(1,f(y)*q2(x,y,theta)/(f(x)*q2(y,x,theta)))
}
```

```{r,echo = FALSE}
fun_scale = function(theta){
  #using fixed scale
  x = 1
  t = 0
  list = c()
  
  #preheat
  for (i in 1:300){
    y = rgamma(1,shape = x/theta,scale = theta)
    u =runif(1,0,1)
    r = r2(x,y,theta)
    if (u<=r){
      x = y
    }
  }
  
  
  for (i in 1:50001){
    if (t > 49) {
      t = 0
      list = c(list,x)
    }
    
    y = rgamma(1,shape = x/theta,scale = theta)
    u =runif(1,0,1)
    r = r2(x,y,theta)
    if (u<=r){
      x = y
      }
    t=t+1
  }
  par(mfrow=c(1,2))
  plot(list,type = "l",xlab = "iteration",ylab = "",
       main = paste("using Gamma(",theta,",x/",theta,")", sep = ""))
  plot(density(list),main = 
         paste("using Gamma(",theta,",x/",theta,")", sep = ""))
  list
}

```

```{r,echo = FALSE}
fun_shape = function(k){
  #using fixed shape
  x = 1
  t = 0
  list = c()
  
  for (i in 1:300){
    y = rgamma(1,shape = k,scale = x/k)
    u =runif(1,0,1)
    r = r1(x,y,k)
    if (u<=r){
      x = y
    }
  }
  
  
  for (i in 1:(50*1000+1)){
    if (t > 49) {
      t = 0
      list = c(list,x)
    }
    
    y = rgamma(1,shape = k,scale = x/k)
    u =runif(1,0,1)
    r = r1(x,y,k)
    if (u<=r){
      x = y
      }
    t=t+1
  }
  par(mfrow=c(1,2))
  plot(list,type = "l",xlab = "iteration",ylab = "",
       main = paste("using Gamma(",k,",x/",k,")", sep = ""))
  plot(density(list),main = 
         paste("using Gamma(",k,",x/",k,")", sep = ""))
  list
}
```

$Gamma(k,\theta)$ has mean $k\theta$

we can draw $y_i$ from $Gamma(k,x_{i-1}/k)$ or $Gamma(x_{i-1}/\theta,\theta)$

generate $y_i$ from $q(y|x_{i-1})$ which is $gamma(k,x_{i-1}/k))$ or $gamma(x_{i-1}/\theta,\theta))$

here is some plot of our output.

```{r,echo = FALSE}
output = data.frame("sample mean"=0,"1/sample mean"=0)
output[1,1] =sqrt(2/1.5)
output[1,2] =sqrt(1.5/2)+1/(2*2)
rownames(output)[1] = "True value"
list = fun_shape(2)
output[2,1] = mean(list)
output[2,2] = mean(1/list)
rownames(output)[2] = "Gamma(2,x/2)"
list = fun_shape(10)
output[3,1] = mean(list)
output[3,2] = mean(1/list)
rownames(output)[3] = "Gamma(10,x/10)"
list = fun_scale(2)
output[4,1] = mean(list)
output[4,2] = mean(1/list)
rownames(output)[4] = "Gamma(x/2,2)"
list = fun_scale(10)
output[5,1] = mean(list)
output[5,2] = mean(1/list)
rownames(output)[5] = "Gamma(x/10,10)"
```

Here is the table showing the mean of the sampele and the mean of 1/sample

```{r,echo=FALSE}
output
```

they are similar enough with each other and it seems all of them provide reasonable estimates.